---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
step_data <- read.csv("activity.csv")
step_data <- subset(step_data, !is.na(steps))
```


## What is mean total number of steps taken per day?
```{r}
steps_per_day <- aggregate(step_data$steps, by=list(Date=step_data$date), FUN=sum)
names(steps_per_day) <- c("Date", "Steps")
hist(steps_per_day$Steps, xlab="Steps", main="Step Frequency")
```

```{r}
mean_step <- as.integer(mean(steps_per_day$Steps))
```
The mean total number of steps taken per day is: `r mean_step`

```{r}
median_step <- as.integer(median(steps_per_day$Steps))
```
And the median total number of steps taken per day is: `r median_step`

## What is the average daily activity pattern?
```{r}
average_steps <- aggregate(step_data$steps, by=list(Date=step_data$date), FUN=mean)
plot(average_steps, type="l", ylab="Average number of steps")
```

The highest average number of steps was on 2012-11-24.

## Imputing missing values
```{r}
step_data <- read.csv("activity.csv")
nas <- subset(step_data, is.na(steps))
na_rows <- nrow(nas)
```
There are `r na_rows` rows containing NA values.

```{r}
step_data$steps <- apply(step_data, 1, function(row) if(is.na(row[1])) { 0 } else { as.numeric(row[1]) })
total_steps <- aggregate(step_data$steps, by=list(Date=step_data$date), FUN=sum)
names(total_steps) <- c("Date", "Steps")
average_steps <- aggregate(total_steps$Steps, by=list(Date=total_steps$Date), FUN=mean)
names(average_steps) <- c("Date", "Average_Steps")
hist(average_steps$Average_Steps)
mean_step <- as.integer(mean(average_steps$Average_Steps))
median_step <-as.integer(median(average_steps$Average_Steps))
```

The new mean steps is `r mean_step`.
The new median steps is `r median_step`.

The new mean and median are different from the earlier estimates. The effect of imputing missing data caused the mean and median to shift towards zero.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
average_steps$wd <- apply(average_steps, 1, function(row) if(weekdays(as.Date(row[1])) == "Saturday" || weekdays(as.Date(row[1])) == "Sunday") { "Weekend" } else { "Weekday" })

```