---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
step_data <- read.csv("activity.csv")
step_data <- subset(step_data, !is.na(steps))
```


## What is mean total number of steps taken per day?
```{r}
steps_per_day <- aggregate(step_data$steps, by=list(Date=step_data$date), FUN=sum)
names(steps_per_day) <- c("Date", "Steps")
hist(steps_per_day$Steps, xlab="Steps", main="Step Frequency")
```

The mean total number of steps taken per day is:
```{r}
mean(steps_per_day$Steps)
```

And the median total number of steps taken per day is:
```{r}
median(steps_per_day$Steps)
```

## What is the average daily activity pattern?
```{r}
average_steps <- aggregate(step_data$steps, by=list(Date=step_data$date), FUN=mean)
plot(average_steps, type="l", ylab="Average number of steps")
```

The highest average number of steps was on 2012-11-24.

## Imputing missing values
```{r}
step_data_with_na <- read.csv("activity.csv")
nas <- step_data_with_na[step_data_with_na$steps == "NA", ]
na_rows <- nrow(nas)
```
There are `r na_rows` rows containing NA values.



## Are there differences in activity patterns between weekdays and weekends?
